---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning=FALSE,
  message=FALSE
)
```

# `usgs.discharge`: A simple package for calculating 30-year USGS discharge percentiles.

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of usgs.discharge from [GitHub](https://github.com/mt-climate-office/usgs-discharge) with:

``` r
# install.packages("devtools")
devtools::install_github("mt-climate-office/usgs-discharge")
```

## How to Use

Get 30 years of data from a USGS gauge, and plot current conditions with USDM drought categories.
```{r example}
library(usgs.discharge)
discharge <-  get_discharge("05014500")
make_climatology_plot(
 discharge,
  "Swiftcurrent Creek at Many Glacier MT",
  "05014500"
)
```

Below is an example of how to both create plots and an `sf` object with current discharge conditions for the entire UMRB. **Note** that this uses a `future` plan with all but 1 of your available cores. I don't know if there is a rate limiting scheme on the USGS API, but be careful using more than ~20 cores. I only tested using 19 cores and didn't run into any issues:

```{r umrb}
library(usgs.discharge)

# Create a future plan for multicore processing. If you don't do this, everything
# will happen sequentially.
future::plan(future::multisession, workers = future::availableCores() -1)
stations <- get_gauges(clip_shp = usgs.discharge::domain)

# only use 5 stations for this example.
stations <- stations |> head(5)

# You can ignore the warning this spits out.
# We only get 3 stations as a result, because this function filters out stations
# that don't have a 30-year observation record.
discharge <- get_discharge_shp(stations)

# Make plots. You can specify an out_dir arguemnt to write the plots out to a folder:
# plots <- furrr::future_pmap(dat, make_climatology_plot, out_dir = "./plots")
plots <- furrr::future_pmap(discharge, make_climatology_plot) 
plots[[2]]

discharge_shp <- calc_discharge_anomalies(discharge)

print(discharge_shp)
```
